Statically Typed String Sanitation Inside a Python

### 2014-08-21 sanitation-psp14/sanitation-psp14-short.tex
updated

\renewcommand{\grammarlabel}[2]{#1\hfill#2}
\begin{figure}[h]
\small
  \begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

@@

\begin{figure}[h]
\small
  \begin{grammar}

@@

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ \alt
      $e(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ | $\strcase{e}{e}{x,y.e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ | $\rcheck{r}{e}{x.e}{e}$

      <$v$> ::= $\lambda x . e$ | $\strin{e}$ | $s$ \hfill source values 
\caption{Syntax of $\lambda_{RS}$.}
\label{fig:glambdas}
\end{grammar}
\end{figure}

@@

Whereas concatenation allows the construction of large strings from smaller strings, \textsc{S-T-Case} allows the decomposition, or elimination, of large strings into smaller strings. Intuitively, this rule ``peels off" the first character, and then performs a specified operation on the first character and the remaining string.

In terms of strings, \textsc{S-T-Case} is essentially the ubiquitious substring operation.

@@

\begin{defn}[Definition of \lhead{r}]
The function $\lhead{r}$ is defined in terms of the structure of $r$:
\begin{itemize}[noitemsep]
\item $\lhead{a r'} = a$ where $a \in \Sigma$
\item $\lhead{\epsilon} = \epsilon$
\item $\lhead{r_1 +  r_2}  = \lhead{r_1} + \lhead{r_2}$
\item $\lhead{r*} = \lhead{r}$
\item $\lhead{.} = a_1 + a_2 + ... + a_n$ for all $a_i \in \Sigma$ where $|\Sigma| = n$.
\end{itemize}
\end{defn}

@@

An alternative to coercion is subtyping for constrained string types, which applies coercion implicitly. 
In practice, subtyping is crucial to the usability of our system due to the nature of the replacement operator.
For simplicity, we do not include this relation; however, subtyping for constrained strings is present in \cite{fulton12,fulton13} and that treatment extends to this system. The implementation discussed in Section \ref{atlang} provides subtyping between constrained string types.

@@

\subsection{Target Language}\label{sec:p}
\begin{figure}[t]
\small
  \begin{grammar}

@@

<$\iota$> ::= $x$  \hfill target terms \alt
             $\lambda x . \iota$ \alt
                    $\iota(\iota)$ \alt
$\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{\iota}{\iota}$ | $\pstrcase{\iota}{\iota}{x,y.\iota}$ \alt
  $\preplace{\iota}{\iota}{\iota}$ | $\tcheck{\iota}{\iota}{\iota}{\iota}$

  <$\dot{v}$> ::= $\lambda x . \iota$ | \tstr{s} | \rx{r} \hfill target values

@@

@@

\begin{figure}[h]\begin{lstlisting}
from atlib import fn, stringin

@@

% HosoyaPierce2002? 
Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Our system is defined within an extensible type system.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML schema.
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
\end{itemize}

@@

We believe that this type system extension serves as a useful basis for web-oriented static analysis;
frameworks and regular expression libraries could be annotated, along with use-sites.
%The use sites of arbitrary strings are usually constrained to a few sections of the application. Assuming
%regular expressions are used for input sanitation, annotations at these use sites could be sufficient input to a sound and complete static analysis.
%
We also believe that extensible programming languages are a promising approach toward incorporating other security
and privacy analyses into programming languages.
%languages might provide developers with a way to incorporate domain-specific security and privacy analyses into their
%daily development activities. In future work, we hope to bolster this hypothesis by extending our approach in this paper
%to more realistic settings and by developing new type extensions addressing security and privacy concerns.
%
\section{Conclusion}

>>>>

\renewcommand{\grammarlabel}[2]{#1\hfill#2}
\begin{figure}[h]
\begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

@@

\begin{figure}[h]
\begin{grammar}

@@

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ e\alt
      $e(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\strcase{e}{e}{x,y.e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ \alt
      $\rcheck{r}{e}{x.e}{e}$

<$v$> ::= $\lambda x . e$ \hfill source values \alt
          $\strin{s}$ \hfill $s \in \Sigma^{*}$

\caption{Syntax of $\lambda_{RS}$.}
\label{fig:glambdas}
\end{grammar}
\end{figure}

@@

Whereas concatenation allows the construction of large strings from smaller strings, \textsc{S-T-Case} allows the decomposition, or elimination, of large strings into smaller strings. Intuitively, this rule ``peels off" the first character, and then performs a specified operation on the first character and the remaining string. In terms of strings, this is essentially the ubiquitious substring operation.

@@

\begin{defn}[Definition of \lhead{r}]
The function $\lhead{r}$ is defined in terms of the structure of $r$:
\begin{itemize}
\item $\lhead{a r'} = a$ where $a \in \Sigma$
\item $\lhead{\epsilon} = \epsilon$
\item $\lhead{r_1 +  r_2}  = \lhead{r_1} + \lhead{r_2}$
\item $\lhead{r*} = \lhead{r}$
\item $\lhead{.} = a_1 + a_2 + ... + a_n$ for all $a_i \in \Sigma$ where $|\Sigma| = n$.
\end{itemize}
\end{defn}

@@

An alternative to coercion is subtyping for constrained string types. For simplicity, we do not include this relation; however, subtyping for constrained strings is present in \cite{fulton13} and that treatment extends to this system.

@@

\subsection{Target Language}\label{sec:p}
\begin{figure}[t]
\begin{grammar}

@@

<$\iota$> ::= $x$  \hfill target terms \alt
             $\lambda x . \iota$ \alt
                    $\iota(\iota)$ \alt
$\tstr{s}$ \alt
  $\rx{r}$ \alt
  $\tconcat{\iota}{\iota}$ \alt
  $\pstrcase{\iota}{\iota}{x,y.\iota}$ \alt
  $\preplace{\iota}{\iota}{\iota}$ \alt
  $\tcheck{\iota}{\iota}{\iota}{\iota}$

<$\dot{v}$> ::= $\lambda x . \iota$ \hfill target values \alt
                $\tstr{s}$ \alt
                $\rx{r}$

@@

In practice, the type of a replacement rarely matches a specification.
Therefore, it is convienant in an implementation to always insert the appropriate
safe coercion. This way, violations of the sanitation protocol manifest as type errors,
but the definition of an adequate replacement is not arduous.
Alternatively, this policy may 
be codified in the type system itself using subtyping \cite{fulton12}.

@@

\begin{figure}[h]\small\begin{lstlisting}
from atlib import fn, stringin

@@

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP,HosoyaPierce2002} checks XML documents against schema using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least three ways:
\begin{itemize}
  \item Our system is defined within an extensible type system.
  \item We demonstrate that regular expression types are applicable to the web security domain, whereas previous work on regular expression types focused on XML schema.
  \item Although our static replacement operation is definable in some languages with regular expression types, we are the first to expose this operation and connect the semantics of regular language replacement with the semantics of string substitution via a type safety and compilation correctness argument.
\end{itemize}

@@

We believe that this type system extension serves as a useful basis for web-oriented static analysis.
The use sites of arbitrary strings are usually constrained to a few sections of the application. Assuming
regular expressions are used for input sanitation, annotations at these use sites could be sufficient input to a sound and complete static analysis.

We also believe that extensible programming languages are a promising approach toward incorporating security
and privacy into programming languages. The Software Industry moves quickly, and developing new languages for 
the particular security and privacy concerns of each new domain is likely a losing strategy. Conversely, extensible
languages might provide developers with a way to incorporate domain-specific security and privacy analyses into their
daily development activities. In future work, we hope to bolster this hypothesis by extending our approach in this paper
to more realistic settings and by developing new type extensions addressing security and privacy concerns.

\section{Conclusion}
